False Positive Probabilities for all Kepler Objects of Interest: \\ \nvalnew\ newly validated planets and \nfpnew\ likely false positives

### 2015-08-27 ms.tex
paper edits and pie chart exploration

\newcommand{\nvalnew}{1000}
\newcommand{\nfp}{500}
\newcommand{\nfpnew}{200}
\newcommand{\nbadphot}{47} %EmptyPhotometry or MissingKOI
\newcommand{\nbadphotFP}{34}
\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{216}
\newcommand{\nbadstellar}{98} %MissingStellar
\newcommand{\nbadsec}{162} %NoWeakSecondary
\newcommand{\nattempted}{8141}
\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{32}
\newcommand{\nbadroche}{52}
\newcommand{\nbadother}{28}
\newcommand{\nbadotherFP}{22}
\newcommand{\nbadotherCAND}{5}
\newcommand{\nbadotherCONF}{1}

\newcommand{\nreliable}{2854}
\newcommand{\nval}{2136}
\newcommand{\nreliableFP}{298}

@@

\vespa\ models an eclipse signal as a simple trapezoid, parametrized
by depth $\delta$, total duration $T$, and shape parameter $T / \tau$,
where $\tau$ is the ``ingress/egress'' duration (such that a
completely V-shaped transit has $T/\tau = 2$).  For the transit signal
being evaluated, the joint posterior probability density function
(PDF) of these shape parameters is sampled with Markov Chain Monte
Carlo (MCMC), using the
\texttt{emcee} sampler \citep{emcee}.  This allows the likelihood for
each hypothesis to be determined by simulating a physically realistic
population of the hypothesized astrophysical scenario and using this
population to define the PDF for the trapezoidal parameters under the
hypothesis.  The likelihood is then
\begin{equation}
  \eqlabel{lhood}
  \mathcal L_i = \displaystyle \int p_\mathrm{sig}\left(\bvec{\theta}\right)
                                    p_i\left(\bvec{\theta}\right)\,d\bvec{\theta},
\end{equation}
where $\bvec{\theta}$ is the vector of trapezoidal shape parameters,
$p_\mathrm{sig}$ is the posterior PDF of the signal, and $p_i$ is the
PDF for the parameters under hypothesis $i$.\footnote{$\mathcal L_i$
  may be seen to be the ``evidence'' or ``marginalized likelihood'' of
  the trapezoidal model under hypothesis $i$, with $p_\mathrm{sig}$
  being the likelihood and $p_i$ being the prior, integrated over the
  $\bvec{\theta}$ parameter space.  But for clarity, and for
  continuity with previous publication, we continue to call $\mathcal
  L_i$ the ``likelihood'' for hypothesis $i$.}

@@

We emphasize that stellar parameter estimation is not the central goal
of this work, nor are the FPP results very sensitive to the exact
estimated stellar properties.  The exception to this would be if the
stellar density estimate is significantly mis-estimated, which would
be the case if a star is not properly identified as evolved.  However, we
note that of the 743 KOIs with host stars $>$2\,$R_{\odot}$ at the 
NExScI Archive, 515 of them already have FALSE POSITIVE designations;
additionally, \citet{Sliski:2014} find a large false positive rate for
KOIs with evolved host stars.  Given all of these considerations, we
believe that potential systematic issues with stellar property
determinations in various corners of parameter space do not affect the
main results, which are the false positive probabilities.

@@

Of the \ntotal\ KOIs in the cumulative table at the NExScI Exoplanet
Archive, \vespa\ successfully calculated the FPP for \ncalc.  These
results are listed in \tab{fpp};  \sect{failures} contains detailed
explanations of the different failure modes.

@@

Another important constraint used in the FPP calculation is the
allowed sky area inside which a blended false positive may live.  As
described in \sect{data}, this value is taken to be three times the
uncertainty on the fitted centroid position from the pixel-level data.
However, many KOIs have already been identified to be blended binaries
displaced from the target star.  Some of these are found by detecting
significant centroid offsets---in these cases, treating the confusion
radius simply as the uncertainty in the centroid position will clearly
give a misleading result.  Other displaced false positives have been
identified as originating from displaced stars by finding KOIs with
matching ephemerides \citep{Coughlin:2014}, and often the ``parents''
of these signals are outside the pixel masks, and so unable to be
detected via centroid-measuring methods.  In these cases as well, the
\vespa\ assumptions break down, and the FPP calculations will not be
valid.

To summarize, the results presented in \tab{fpp} are
strictly reliable only for KOIs \emph{that have already strongly
  passed the other Kepler vetting tests.}  The first cut for this is
the KOI disposition: FALSE POSITIVE indicates failure of one or more
of these tests (and thus probable invalidity of the
\vespa\ calculation).  However, because of the generally permissive
philosophy of the \kepler\ dispositioning, not all CANDIDATE KOIs
receive the same level of vetting.  That is, something is not
identified as a FALSE POSITIVE \emph{unless there is positive
  confirmation of false positive status}.  In particular, when
pixel-level analysis fails to determine whether the signal is indeed
coming from the target star because of low SNR, such a KOI will still
receive CANDIDATE disposition.  Therefore, the \vespa-calculated FPP
may be considered reliable only when a KOI is designated a CANDIDATE
(or CONFIRMED) and has large enough SNR to enable secure positional
determination.

To enable convenience of interpretation, \tab{fpp} thus contains both
the current KOI disposition and the results of the positional
probability calculations of \citet{Bryson:2015}.  \fig{fppall} shows
the calculated FPPs for all KOIs not dispositioned FALSE POSITIVE that
have $>$99\% probability to be located on the target star (and for
which the positional calculations are reliable; that is, having a
``reliability score'' > 0.4).

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_summary_all.pdf}
\end{center}
\caption{False positive probabilities of all KOIs for which we
  consider the \vespa\ calculations to be reliable (floored at
  $10^{-4}$ for visualization purposes).  This includes KOIs that are
  not dispositioned as FALSE POSITIVE on the NExScI archive, are
  considered to be reliably located on the target star ($ {\rm Pr} >
  0.99$, with ``score'' $>$ 0.4) according to pixel-level analysis
  \citep{Bryson:2015}, and for which $\mathcal L_{\rm tot} > 10^{-3}$,
  indicating that the hypotheses tested by \vespa\ are not all poor
  explanations of the data.  Of these \nreliable\ KOIs, \nval\ have
  FPPs less than 1\%.  Noteably, \nreliableFP\ are likely false
  positives (FPP $> 0.5$), consistent with the \citet{Morton:2011} and
  \citet{Fressin:2013} \emph{a priori} estimates of the overall
  \kepler\ candidate false positive rate.  \figlabel{fppall}}
\end{figure*}

@@

>>>>

\newcommand{\nfp}{500}
\newcommand{\nfpnew}{200}
\newcommand{\nbadphot}{47} %EmptyPhotometry or MissingKOI
\newcommand{\nbadphotFP}{34}
\newcommand{\nbadphotCAND}{13}
\newcommand{\nbadrowemcmc}{216}
\newcommand{\nbadstellar}{98} %MissingStellar
\newcommand{\nbadsec}{162} %NoWeakSecondary
\newcommand{\nattempted}{8141}
\newcommand{\ntryfail}{112}
\newcommand{\nbadtrapfit}{32}
\newcommand{\nbadroche}{52}
\newcommand{\nbadother}{28}
\newcommand{\nbadotherFP}{22}
\newcommand{\nbadotherCAND}{5}
\newcommand{\nbadotherCONF}{1}

\newcommand{\nreliable}{2752}
\newcommand{\nval}{2069}
\newcommand{\nreliableFP}{292}
\newcommand{\nvalnew}{1368}
\newcommand{\ncurval}{1001}

@@

\vespa\ models an eclipse signal as a simple trapezoid, parametrized
by depth $\delta$, total duration $T$, and shape parameter $T / \tau$,
where $\tau$ is the ``ingress/egress'' duration (such that a
completely V-shaped transit has $T/\tau = 2$).  For the transit signal
being evaluated, the joint posterior probability density function
(PDF) of these shape parameters is sampled with Markov Chain Monte
Carlo (MCMC), using the
\texttt{emcee} sampler \citep{emcee}.  This allows the likelihood for
each hypothesis to be determined by simulating a physically realistic
population of the hypothesized astrophysical scenario and using this
population to define the PDF for the trapezoidal parameters under the
hypothesis.  The likelihood is then
\begin{equation}
  \eqlabel{lhood}
  \mathcal L_i = \displaystyle \int p_\mathrm{sig}\left(\bvec{\theta}\right)
                                    p_i\left(\bvec{\theta}\right)\,d\bvec{\theta},
\end{equation}
where $\bvec{\theta}$ is the vector of trapezoidal shape parameters,
$p_\mathrm{sig}$ is the posterior PDF of the signal, and $p_i$ is the
PDF for the parameters under hypothesis $i$.\footnote{$\mathcal L_i$
  may be seen to be the ``evidence'' or ``marginalized likelihood'' of
  the trapezoidal model under hypothesis $i$, with $p_\mathrm{sig}$
  being the likelihood and $p_i$ being the prior, integrated over the
  $\bvec{\theta}$ parameter space.  But for clarity, and for
  continuity with previous publication, we continue to call $\mathcal
  L_i$ the ``likelihood'' for hypothesis $i$.}  The hypotheses
currently supported by \vespa are the following: unblended eclipsing
binary (EB), hierarchical-triple eclipsing binary (HEB),
chance-aligned background/foreground eclipsing binary (BEB), and
transiting planet (Pl).

@@

We emphasize that stellar parameter estimation is not the central goal
of this work, nor are the FPP results very sensitive to the exact
estimated stellar properties.  The exception to this would be if the
stellar density estimate is significantly mis-estimated, which would
be the case if a star is not properly identified as evolved.  However,
we note that of the 743 KOIs with host stars $>$2\,$R_{\odot}$ at the
NExScI Archive, 515 of them already have FALSE POSITIVE designations;
additionally, \citet{Sliski:2014} find a large false positive rate for
KOIs with evolved host stars.  Given all of these considerations, we
believe that potential systematic issues with stellar property
determinations in various corners of parameter space do not strongly
affect the main results, which are the astrophysical false positive
probabilities of thousands of KOIs.

@@

Of the \ntotal\ KOIs in the cumulative table at the NExScI Exoplanet
Archive, \vespa\ successfully calculates the FPP for \ncalc.  These
results are listed in \tab{fpp};  \sect{failures} contains detailed
explanations of the failure reasons for the other \nfail\ KOIs.

@@

However, while the FPP value does only give the \emph{relative}
probability between the false positive and planet scenarios, there is
a way to get a sense of whether \emph{all} of the models might be bad
fits to the data (and thus get a hint that perhaps the signal is due
to stellar variability or a false alarm).  Recall that FPP is
calculated as
\begin{equation}
\eqlabel{fpp}
{\rm FPP} = \frac{\displaystyle\sum_{i \neq {\rm Pl}}
  L_i}{\displaystyle \sum_i L_i}
\end{equation}
where $L_i$ = $\mathcal L_i \pi_i$ and $i$ runs over all the models,
with ``Pl'' representing the planet scenario.  If, for example, the
shape of the transit is such that \emph{none} of the models gives any
kind of decent fit to its trapezoidal shape, then $L_{\rm tot} =
\sum_i L_i$ will be very small.  It appears that a value of $10^{-3}$
is a reasonable threshold below which to consider a particular
calculation unreliable.  To set this threshold we inspected the
handful of systems that have low FPPs and $L_{\rm tot} < 10^{-3}$
(some go as low as $10^{-10}$ or lower), and find that these very
often have unrealistically long durations for a transit signal.  Such
a case could easily give the planet hypothesis the \emph{largest}
likelihood (resulting in a low FPP), not because it is intrinsically
likely, but because everything else is just much worse.

Another important constraint used in the FPP calculation is the
allowed sky area inside which a blended false positive may live.  As
described in \sect{data}, this value is taken to be three times the
uncertainty on the fitted centroid position from the pixel-level data.
However, many KOIs have already been identified to be blended binary
false positives displaced from the target star.  Some of these are
found by detecting significant centroid offsets in the pixel-level
data---in these cases, \vespa\ treating the confusion radius simply as
the uncertainty in the centroid position will clearly give a
misleading result.  Other displaced false positives have been
identified as originating from displaced stars by finding KOIs with
matching ephemerides \citep{Coughlin:2014}, and often the ``parents''
of these signals are outside the pixel masks, and so unable to be
detected via centroid-measuring methods.  In these cases as well, the
\vespa\ assumptions break down, and the FPP calculations will not be
valid.

To summarize, the results presented in \tab{fpp} are strictly reliable
only for KOIs \emph{that have already strongly passed the other Kepler
  vetting tests, and that are not indicated to be clearly poor fits to
  all the proposed hypotheses}.  The first cut for this is the KOI
disposition: FALSE POSITIVE indicates failure of one or more of these
tests (and thus probable invalidity of the \vespa\ calculation).
However, because of the generally permissive philosophy of the
\kepler\ dispositioning, not all CANDIDATE KOIs receive the same level
of vetting.  That is, something is not identified as a FALSE POSITIVE
\emph{unless there is positive confirmation of false positive status}.
In particular, when pixel-level analysis fails to determine whether
the signal is indeed coming from the target star because of low SNR,
such a KOI will still receive CANDIDATE disposition.  Therefore, the
\vespa-calculated FPP may be considered most reliable only when a KOI
is designated a CANDIDATE (or CONFIRMED) and has large enough SNR to
enable secure positional determination.

To enable interpretation, \tab{fpp} thus contains both the current KOI
disposition and the results of the positional probability calculations
of \citet{Bryson:2015}.  \Fig{fppall} shows FPPs for all KOIs passing
the following criteria:
\begin{itemize}
\item Dispositioned CANDIDATE or CONFIRMED at the NExScI archive
\item $> 0.99$ probability to be on the target star, according
  to \citet{Bryson:2015}, along with a positional probability
  ``score'' $> 0.4$ (indicating a reliable result).
\item $L_{\rm tot} > 1e-3$, indicating that not all the hypotheses are
  poor fits to the transit signal.
\end{itemize}

These selections leave \nreliable\ KOIs for which the \vespa\ results
can be considered most reliable.  Of these, \nval\ have ${\rm FPP} <
0.01$, and are thus validated at the 99\% level.

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_summary_all.pdf}
\end{center}
\caption{False positive probabilities of all KOIs for which we
  consider the \vespa\ calculations to be reliable (floored at
  $10^{-4}$ for visualization purposes).  This includes KOIs that are
  not dispositioned as FALSE POSITIVE on the NExScI archive, are
  considered to be reliably located on the target star ($ {\rm Pr} >
  0.99$, with ``score'' $>$ 0.4) according to pixel-level analysis
  \citep{Bryson:2015}, and for which $L_{\rm tot} > 10^{-3}$,
  indicating that the hypotheses tested by \vespa\ are not all poor
  explanations of the data.  Of these \nreliable\ KOIs, \nval\ have
  FPPs less than 1\%.  Noteably, \nreliableFP\ are likely false
  positives (FPP $> 0.5$), consistent with the \citet{Morton:2011} and
  \citet{Fressin:2013} \emph{a priori} estimates of the overall
  \kepler\ candidate false positive rate.  \figlabel{fppall}}
\end{figure*}
